---
title: "Investigating a Hospital-Specific Report (HSR)"
subtitle: "Analysis ideas for gaining insight into HRRP program results."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hsr-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

_Work in progress_

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

_**Note**: CMS changed the format of Hospital-Specific Reports (HSRs) for FY2026 (see [here](https://qualitynet.cms.gov/inpatient/hrrp/reports#tab2)). The current HSR functions support Excel-based formats through FY2025. However, analysis strategies are still relevant._

As part of the [Hospital Readmissions Reduction Program (HRRP)](https://www.cms.gov/medicare/payment/prospective-payment-systems/acute-inpatient-pps/hospital-readmissions-reduction-program-hrrp), the [Centers for Medicare & Medicaid Services (CMS)](https://www.cms.gov/) provides a detailed, annual program summary report (called the [Hospital-Specific Report (HSR)](https://qualitynet.cms.gov/inpatient/hrrp/reports)) to hospitals that includes details on the penalty calculation for the upcoming fiscal year, such as discharge-level data, dually-eligible discharge lists, cohort-level rollup, and the penalty amount. There is a defined [review and correction period](https://qualitynet.cms.gov/inpatient/hrrp/resources#tab1:~:text=FY%202026%20Hospital%20Readmissions%20Reduction%20Program%20Key%20Dates%20(08/11/25)) in which hospitals can use these reports to ensure the penalty being enforced by CMS is accurate. It occurs approximately 1 month before the new fiscal year, thus it is a time-critical event (see historical date ranges below for reference with the built-in package datasets):

```{r}
# Extract date ranges
readmit::hrrp_keydates |>
  dplyr::select(
    ProgramYear,
    dplyr::matches("^(Payment|Review)")
  ) |>
  dplyr::distinct()
```

The report file itself (through FY2025) is a large, multi-tab Microsoft Excel document where the structured part of the data is ambiguously placed throughout, thus we need tools to parse it out into a usable format. That is what some functions in the `readmit` package are for. In this article, we go through the tools that are available, what they do, and then provide some strategies/approaches for how hospitals can use these tools to analyze their own HSR's to gain deeper insight into HRRP results and readmissions more broadly.

# The Toolbox

First, we'll start by taking a look at what relevant functions are available to us, what they do, and how to use them. For our purposes, these are all of the functions prefixed like `hsr_*`. We'll do this roughly in order of how the report is laid out, and how the HRRP results roll up.

```{r}
library(readmit)
```

## 0. Mock Reports

As the developer of this package, I don't have access to hospitals' actual HSR's, as they contain senstivie patient information (i.e., [PHI](https://cphs.berkeley.edu/hipaa/hipaa18.html)) and thus are not publicly available. So what we have to work with are _mock_ reports that [CMS provides](https://qualitynet.cms.gov/inpatient/hrrp/reports#tab3) to the public that are meant to mimick the format a hospital can expect their report to be in. It just includes fake data.

_**Note**: CMS changed the format of Hospital-Specific Reports (HSRs) for FY2026 (see [here](https://qualitynet.cms.gov/inpatient/hrrp/reports#tab2)). The current HSR functions support Excel-based formats through FY2025._

Nevertheless, these provide a useful playground to analyze the mechanics of the program. We'll start by finding a report with the `hsr_mock_reports()` function:

* Using no arguments lists the various mock files included in the package

```{r}
hsr_mock_reports()
```

* Entering one of the listed file names will return the complete path to the file in the packages location on your computer

```{r}
my_report <- hsr_mock_reports("FY2025_HRRP_MockHSR.xlsx")
my_report
```

Now we can use that HSR path with other package functions. Of course, you would just point to your own HSR when analyzing your hospital's reports.

## 1. Program Summary

We'll start with high level program results. Ultimately, all of the moving parts in the HRRP roll up into a single number: the penalty amount applied to your hospital. This is typically the first table in your report. We can parse it out of the report with the `hsr_payment_summary()` function:

```{r}
my_payment_summary <- hsr_payment_summary(my_report)
my_payment_summary
```

We now have this information in a data frame that we can manipulate as needed:

```{r}
my_payment_summary |>
  tidyr::pivot_longer(dplyr::everything())
```

There are also helper functions to extract specific components from this table.

```{r}
hsr_payment_penalty(my_report)
hsr_dual_proportion(my_report)
hsr_peer_group(my_report)
```

See `?hsr_payment_summary` for all of them.

## 2. Cohort Summary {#cohortsummary}

The overall payment penalty a hospital receives is a weighted average of penalities applied to the individual cohorts. These details are typically in the second table (tab) of the HSR, which we can import with `hsr_cohort_summary()`:

```{r}
cohort_summary <- hsr_cohort_summary(my_report)
cohort_summary
```

We can then, for example, reconcile the overall penalty amount based on what is in this table:

```{r}
cohort_summary |>

  # Filter to cohorts with a penalty
  dplyr::filter(`Penalty Indicator (Yes/No) [h]` == "Yes") |>
  
  # Compute the contribution of each penalized cohort
  dplyr::mutate(
    Contribution = 
      `Ratio of DRG Payments Per Measure to Total Payments [i]` *
      (
        `Excess Readmission Ratio (ERR) [f]` - 
          `Peer Group Median ERR [g]`
      )
  ) |>
  
  # Roll up into final calculation
  dplyr::summarize(
    Penalty = sum(Contribution) * hsr_neutrality_modifier(my_report)
  )
```

What we did here was:

1. Find cohorts who received a penalty _who were eligible_
    + A cohort must have at least 25 discharges to be eligible
    + Then, the excess readmission ratio (ERR) must be greater than the assigned peer group's median ERR
2. Find the difference between the hospital's ERR compared to the peer group median ERR
3. Multiply that by the ratio of DRG payments for that cohort
    + This is a measure of the volume of patients with this condition are treated at the hospital
4. Sum those contributions over each cohort
5. Multiply that by the neutrality modifier in the `hsr_payment_summary()` table

That's how the penalty is computed, from cohort summary level.

## 3. Discharges

_Note: Because we are using mock reports, the dates in these files are erroneous and thus R doesn't interpret them as dates. However, your hospital report has real dates and thus R should automatically parse them as such._

The HSR also contains discharge-level data on the individual patients that actually contributed to the program. There is a separate table/tab for each of the cohorts. We can use the `hsr_discharges()` function to import them for a specified cohort:

```{r}
hsr_discharges(my_report, "COPD")
```

We get some patient identifying information, including the specific dates associated with the index and readmission hospitalizations, whether or not the readmission occurred at the same hospital, diagnosis codes, etc., which is all very valuable information that we can explore to gain insights from (and what we'll do [later](#analysisstrategies)).

There are also options available in the function to refine the result:

### Eligible Cases

The `eligible_only` argument can be used to only included discharges that were actually included in HRRP evaluation:

```{r}
hsr_discharges(
  file = my_report, 
  cohort = "COPD",
  eligible_only = TRUE
)
```

Notice that this row count matches what was reported as the COPD denominator in the `cohort_summary`:

```{r}
cohort_summary |>
  dplyr::filter(`Measure [a]` == "COPD") |>
  dplyr::pull(`Number of Eligible Discharges [b]`)
```

### Risk Factors

Also included in these tables are the indicators of risk factors that are used in the statistical models to estimate individual adjusted readmission risks. We can use the `risk_factors` argument to extract those for each patient:

```{r}
hsr_discharges(
  file = my_report, 
  cohort = "COPD",
  eligible_only = TRUE,
  risk_factors = TRUE,
  discharge_phi = FALSE
)
```

This data can then be explored further to understand risk factor prevalence and how that relates to model weights, etc. (again, covered [later](#analysisstrategies)). Notice the `discharge_phi` argument was used to prevent the date information from being returned.

## 4. Model Coefficients

The ERR is calculated based on an aggregated roll up of individual adjusted readmission risks derived from a random-intercept logistic regression model. The first row in the discharge table contains the coefficients for this model. We can use the `hsr_coefficients()` function to extract them:

```{r}
copd_model <- hsr_coefficients(my_report, "COPD")
copd_model
```

This allows us to do things like assess the relative contribution of risk factors to the estimated readmission rates or use the risk factor dataset above to compute individual level readmission risks.

### Intercept Terms

The _predicted_ and _expected_ readmission rates only differ in the intercept terms applied to the prediction (thus it is a constant shift for all patients). We can see these at the end of this data frame:

```{r}
copd_model |> tail()
```

## 5. Readmission Risks

We could take the risk factor output from `hsr_discharges()` combined with the coefficients from `hsr_coefficients()` and reconcile each patient's _predicted_ and _expected_ readmission risk. But the `hsr_readmission_risks()` function can do all of that for us:

```{r}
risks <- hsr_readmission_risks(my_report, "COPD")
risks
```

This just takes a weighted-sum of the risk factors and coefficients, adds the corresponding intercept, and then maps it to a probability through the logistic function.

The cohort-level _predicted_ and _expected_ readmission rates are just the averages of these columns across all eligible patients:

```{r}
risks |>
  dplyr::summarize(
    Discharges = dplyr::n(),
    Predicted = mean(Predicted),
    Expected = mean(Expected),
    ERR = Predicted / Expected
  )
```

Again, looking at our cohort summary, we can see these match:

```{r}
cohort_summary |> 
  dplyr::filter(
    `Measure [a]` == "COPD"
  ) |>
  dplyr::select(
    Discharges = `Number of Eligible Discharges [b]`,
    Predicted = `Predicted Readmission Rate [d]`,
    Expected = `Expected Readmission Rate [e]`,
    ERR = `Excess Readmission Ratio (ERR) [f]`
  )
```

## 6. Dual Stays

CMS puts hospitals into peer groups based on the relative proportion of Medicare patients who are also eligible for Medicaid. This is a measure of socioeconomic status for the hospital population so hospitals are being compared only against other hospitals that are similar (in this regard). These aggregated quantities were found in `hsr_payment_summary()` result:

```{r}
my_payment_summary
```

The `hsr_dual_stays()` function extracts the discharge-level data corresponding to the numerator of the ratio:

```{r}
hsr_dual_stays(my_report)
```

We can see that the row count of this discharge level data matches the first number in the preceding table.

# Analysis Strategies {#analysisstrategies}

In this section we'll go through a collection data analyses that can be conducted, using functions in `readmit` as support, to validate HSR calculations and/or to gain deeper insights into HRRP results.

## 1. Validating the Penalty Calculation

We [previously calculated](#cohortsummary) the payment penalty starting from the cohort-level results. However, as an initial validation step, it is important to go through the mechanics of reconciling the penalty calculation from the discharge-level data to ensure comprehension of how it works. 

_Note: We'll go the slower, more tedious way to do this in the steps below in order to capture all intermediate details for understanding, but will callout where certain steps can be more efficient._

Let's extract the `cohort` strings we need to plug into various function arguments:

```{r}
cohorts <- setdiff(names(hrrp_cohort_inclusion), "ProgramYear")
cohorts
```

### i. Extract discharges

The first thing we need to is extract the set of discharges (i.e., the _denominator_) that contribute to program for each cohort. To do this, we'll iterate through the different cohorts and sequentially use the `hsr_discharges()` function to get the row identifiers that should be included:

```{r}
eligible_discharges <-
  setdiff(cohorts, "CABG") |>

    # Iterate each cohort
    purrr::map_df(

      # Import eligible discharges
      ~hsr_discharges(
        file = my_report,
        cohort = .x,
        eligible_only = TRUE
      ) |>

      # Keep the row identifiers
      dplyr::select(`ID Number`) |>

      # Add cohort identifier
      tibble::add_column(Cohort = .x)
    )
eligible_discharges
```

We can check the counts of these:

```{r}
table(eligible_discharges$Cohort)
```

You can validate that we obtained the correct counts by looking at the [cohort summary](#cohortsummary) we previously created.

```{r}
cohort_summary
```

One caveat was that we already knew there weren't any `CABG` discharges (via the `NA` in the `cohort_summary` table), so we pre-excluded this from our cohort list we iterated through (as it would have caused an error otherwise).

### ii. Extract risk factors

Next we need to extract the sets of risk factors for cohort that go into the readmission risk model. We can again do this by iterating through the `cohort` list with `hsr_discharges`, but extracting the risk factors as well with `risk_factors=TRUE`:

```{r}
setdiff(cohorts, "CABG") |>

    # Iterate each cohort
    purrr::map(

      # Import eligible discharges
      ~hsr_discharges(
        file = my_report,
        cohort = .x,
        risk_factors = TRUE,
        discharge_phi = FALSE
      )
    )
```

Notice that we get the risk factors for each cohort, but all of the columns are different since each cohort has a different model. So one thing we can do is pivot the data with `tidyr::pivot_longer()` to make a long and narrow data frame so that it can all be binded together.

```{r}
risk_factors <-
  setdiff(cohorts, "CABG") |>

      # Iterate each cohort
      purrr::map_df(

        # Import eligible discharges
        ~hsr_discharges(
          file = my_report,
          cohort = .x,
          risk_factors = TRUE,
          discharge_phi = FALSE
        ) |>
        
        # Send risk factors down the rows
        tidyr::pivot_longer(
          cols = -`ID Number`,
          names_to = "Factor",
          values_to = "Value"
        ) |>
        
        # Indicate cohort
        tibble::add_column(Cohort = .x)
      )
risk_factors
```

We can then merge these (using `dplyr::inner_join()`) with the eligilble discharges we previously identified to get the set of risk factors for each eligible discharge:

```{r}
risk_factors <-
  risk_factors |>

    # Join to get eligible only
    dplyr::inner_join(
      y = eligible_discharges,
      by = 
        c(
          "ID Number",
          "Cohort"
        )
    )
risk_factors
```

Of course, we didn't _have_ to do this step separately from step (i), as we could have just used the `eligible_only` argument simultaneously, and ended up in the same place.

```{r}
risk_factors <-
  setdiff(cohorts, "CABG") |>

      # Iterate each cohort
      purrr::map_df(

        # Import eligible discharges
        ~hsr_discharges(
          file = my_report,
          cohort = .x,
          risk_factors = TRUE,
          discharge_phi = FALSE,
          eligible_only = TRUE
        ) |>
        
        # Send risk factors down the rows
        tidyr::pivot_longer(
          cols = -`ID Number`,
          names_to = "Factor",
          values_to = "Value"
        ) |>
        
        # Indicate cohort
        tibble::add_column(Cohort = .x)
      )
risk_factors
```

### iii. Compute individual readmission risks

The _predicted_ and _expected_ readmission rates are computed for each discharge by plugging in each patient's set of risk factors into risk models developed by CMS. We can use the `hsr_coefficients()` function to extract these:

```{r}
model_weights <-
  cohorts |>

  # Iterate each cohort
  purrr::map_df(

    # Import eligible discharges
    ~hsr_coefficients(
      file = my_report,
      cohort = .x
    ) |>
    
    # Indicate cohort
    tibble::add_column(Cohort = .x)
  )
model_weights
```

To use these in our calculations, we need to attach the model weights for each cohort to our current `risk_factors` dataset using `dplyr::inner_join()`:

```{r}
risk_factors <-
  risk_factors |>

    # Join to get weights
    dplyr::inner_join(
      y = 
        model_weights |> 
          
        # Rename the column
        dplyr::rename(
          Weight = Value
        ),
      by = 
        c(
          "Cohort",
          "Factor"
        )
    )
risk_factors
```

#### Compute the linear predictor (weighted-sum)

These are [logistic regression](https://en.wikipedia.org/wiki/Logistic_regression) models, so to convert to a risk estimate, we need to take the weighted-sum of the factor weight with the risk factor value for each discharge.

```{r}
linear_predictors <-
  risk_factors |>

    # Compute weighted sum
    dplyr::summarize(
      LP = sum(Weight * Value),
      .by = 
        c(
          Cohort,
          `ID Number`
        )
    )
linear_predictors
```

You can think of this as the "risk-adjustment" part, where we've adjusted each patient's readmission risk based on their own clinical history. These are on the [logit](https://en.wikipedia.org/wiki/Logit) scale (so not yet probability/risk estimates), but it what we assume to be linearly related to the outcome, thus we call it the "linear predictor". But we are still missing something: **the intercept terms**.

#### Add in the intercepts

The _predicted_ and _expected_ readmission risks are derived from the same risk-adjusted model, the only difference being in the intercept term that is added to complete the linear predictor. We can see those in our model weight list:

```{r}
model_weights |>
  dplyr::filter(
    stringr::str_detect(
      Factor,
      pattern = "EFFECT$"
    )
  )
```

* The `AVG_EFFECT` corresponds to the risk-shift (intercept) associated with being treated at the "average" hospital
* The `HOSP_EFFECT` corresponds to the risk-shift (intercept) associated with being treated at _your_ hospital

These are estimated by CMS using your discharge lists. They are done so _after_ accounting for all of the risk factors, so by comparing them we get a measure of how much more or less likely a patient is to be readmitted at your hospital versus the average hospital, after risk-adjustment.

We can add each one into our current `linear_predictors` to get the complete linear predictors for the predicted and expected readmission rates:

```{r}
linear_predictors <-
  linear_predictors |>

    # Join to get intercepts
    dplyr::inner_join(
      y = 
        model_weights |>

        # Filter to intercepts
        dplyr::filter(
          stringr::str_detect(
            Factor,
            pattern = "_EFFECT$"
          )
        ) |>
        
        # Send over columns
        tidyr::pivot_wider(
          names_from = Factor,
          values_from = Value
        ),
      by = "Cohort"
    ) |>
    
    # Finish LP's
    dplyr::mutate(
      LP_Predicted = LP + HOSP_EFFECT,
      LP_Expected = LP + AVG_EFFECT
    )
linear_predictors
```

#### Transform to probability scale

The last thing we need to do is transform the linear predictors to the probabilty scale using the [logistic function](https://en.wikipedia.org/wiki/Logistic_function). 

```{r}
readmission_risks <-
  linear_predictors |>

    # Transform LP's
    dplyr::mutate(
      dplyr::across(
        c(LP_Predicted, LP_Expected),
        \(x) 1 / (1 + exp(-x))
      )
    ) |>
    
    # Keep a few columns
    dplyr::select(
      `ID Number`,
      Predicted = LP_Predicted,
      Expected = LP_Expected,
      Cohort
    )
readmission_risks
```

Now we have the _predicted_ and _expected_ readmission risk for each patients.

#### Do it easier

We went through all of that computation to see how it works to derive the readmission risks starting with the discharges, but we can get this automatically by using the `hsr_readmission_risks()` function:

```{r}
readmission_risks <-
  setdiff(cohorts, "CABG") |>

    # Iterate each cohort
    purrr::map_df(

      # Import eligible discharges
      ~hsr_readmission_risks(
        file = my_report,
        cohort = .x
      ) |>
      
      # Indicate cohort
      tibble::add_column(Cohort = .x)
    )
readmission_risks
```

This function extracts eligible discharges and computes the readmission risk, so it captures everything we just did in steps i-iii.

### iv. Compute cohort-level results

## Key Observations

* Note that we don't explcitly use readmission indicators in the calculation. It is all baked into the hospital effect.

# Exploring risk distributions

* Model risk factors
* Risk factor prevalence
* Risk factor prevalence by model weight (scatterplot)

# Diagnosis Comparisons

# Outside Hospital Readmissions

```{r setup}
library(readmit)
hsr_mock_reports()
```
